### 2023-04-04 outputs/paper/paper.qmd
Introduction and Abstract for initial submission outline

---
author: "Sofia Sellitto"
date: "April 21, 2023"
date-format: long
format: 
  pdf: 
    documentclass: article 
    geometry: margin = 1in 
    abstract: "This paper follows an examination of the spatial distribution of educational facilities in Canada and whether a relationship exists between facility location and Province/municipality wealth. The research aims to identify any significant patterns and trends in the distribution of educational facilities, particularly in relation to province and municipality wealth. I have found that _________findings________. The findings of this study aim to contribute to a deeper understanding of educational access and inequality in Canada to inform future research and policies aimed at addressing these issues."

toc: true 
toc-depth: 4

bibliography: references.bib
---

# Introduction

P1: Get a source on education facilities in Canada

In America, unequal access to education is a widely recognized issue that appears to particularly plague impoverished districts and States. Multiple factors contribute to this problem, including the school's funding, whether it is a public or private institution, the student-to-teacher ratio, distance from the school, geographic location, familial income, and racial/ethnic background. This disparity in educational access is a deeply structural issue that spatially affects low-income states and districts.

P2: More context and the gap my work fills

Educational funding in America is primarily Local with the Federal government only contributing approximately 9% of their total budget (cite Atlantic). Because of this, high-poverty districts are far less able to contribute quality budgets for schooling, resulting in them spending 15.6% less per student than low-poverty districts (cite Atlantic). Lower-income families residing in impoverished districts are particularly affected by this inequality, as it has the potential to impact their children's future access to higher education and potentially, their socioeconomic status. (socio-economic status). The acknowledgement of unequal access to education in America is and has been widely understood in discourse for quite some time. However, a nation which appears to receive little attention in regard to this topic of study is Canada.

P3: Introduce the dataset and the purpose of the paper and the research question

In this paper, I aim to investigate the spatial distribution of educational facilities in Canada to address the two following research claims: (1) What is the distribution of educational facilities across Canada & (2) Do they differ spacially and numerical in accordance with the wealth of the Province & municipality wealth. Using an Open Source Data made available by Statistics Canada, I have conducted an exploration of this topic and found that \_\_\_\_\_\_\_\_\_\_\_\_\_\_\_INSERT FINDINGS HERE\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_.

P4: Provide a layout for the paper (what was done, what was found and why its important)

This paper will first begin by \_\_\_\_\_\_\_\_TALK ABOUT DATA SECTION\_\_\_\_\_\_\_\_\_\_\_\_. I will then go onto present \_\_\_\_\_\_\_\_\_\_TALK ABOUT FIGURES AND TABLE\_\_\_\_\_\_. Next I discuss the impacts of my findings on \_\_\_\_\_\_\_\_\_\_\_\_TALK ABOUT DISCUSSION/ LITERATURE\_\_\_\_\_\_\_. Where I commence by discussion implication of my study on future research and push for policy changes.

# Data {#sec-data}

# Model

# Results

# Discussion

## Limitations

##  Next Steps

\newpage

\appendix

# Appendix {.unnumbered}

# Additional details

\newpage

# References